### 2022-03-02 ch1.tex
Updates

A critical turning point in the study of the flow information was put forth by renowned Sociologist Paul Lazarsfeld \citeyearpar{lazarsfeldPeopleChoice1944}. Lazarsfeld and others theorized that the media did not directly influence the masses, but that a "Two-Step Flow" was occurring where "opinion leaders" interpreted messages from mass media and then, in turn, promoted their opinions to the masses \citep{katzPersonalInfluencePart1955}. The opinion leaders interpret, explain, and diffuse contents to "opinion followers". From a networks
perspective, "people rarely act on mass-media information unless it is also transmitted thro\citep[p. 1374]{granovetterStrengthWeakTies1973}. Lazarsfeld's first study focused on politics \citeyearpar{lazarsfeldPeopleChoice1944} but follow up studies from \citep{katzPersonalInfluencePart1955} showed that this model matched information flows beyond the political realm. In contrast to the one-step flow, where media directly influences the masses, Lazarsfeld's theory remains in prominence today. While many scholars assumed that
modern micro-targeted marketing strategies would provide evidence for a one-step flow perspective \citep{bennettOneStepFlowCommunication2006}, many scholars find that "opinion leaders" (often influencers and celebrities) still create a two-step flow of information transmission \citep{choi15, hilbertOneStepTwo2017}.

@@

\subsubsection{Uses and Gratification Theory}

@@

\subsubsection{Social Norms}

@@

\subsection{Broader Impact}

@@

This dissertation also contributes to computational social science &
critical big data studies. I base much of my dissertation around Google
Search Trends, which are relatively underutilized in the social sciences
compared to health sciences and business. Not only does this project
introduce Google Search Trends in new ways, but I perform explicit tests
of how this data should be used in the social sciences, providing the
groundwork for this exciting source of big data for future computational
social scientists. Because my research takes a critical big data studies
approach, I take into account the potential pitfalls with any source of
big data (McFarland and McFarland 2015) and attempt to explicitly answer
questions posed by this field, such as "what are the behavioral
processes that lead to macro-level outcomes" proposed by \citet{breigerScaling2015}.

@@

\subsection{Three Empirical Articles}

@@

\subsubsection{Article 1: Construct Validity and Correspondence of Google Trends}
% TODO Double check name
% TODO link to article  
I will first investigate how Google Search Trends can be operationalized
as cultural attitudes, disease prevalence and voting behaviors through
comparing trends data to observed valid indicators of these cases.
Specifically, I will test cultural indicators from the NORC General
Social Survey, United States county-level suicide rates from National
Center for Health Statistics, US case rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the
American National Election Survey against selected Google Trends terms
and analyze the data using correlation methods. The aim of this article
is to assess the construct validity and will contribute to critical big
data studies and sociological methodology.
% TODO ADD IN ARTICLE 1 ABSTRACT

\subsubsection{Article 2: Information-Seeking Behaviors for Information on Covid-19 Vaccinations}
% TODO Double check name
% TODO link to article  
The second article in my dissertation investigates the question "what
are the behavioral processes that lead to macro-level outcomes" (in this
case, Google Trends results) proposed by \citet{breigerScaling2015}. Knowing that
Google trends data only encompasses a small portion of the
information-seeking done by modern humans, it is important to
investigate what leads individuals to search for a question online using
a search engine (contributing to macro-level trends data) versus more
traditional means of information gathering, such as network activation
(asking someone in their social network for informational support). To
investigate this, I use the case of receiving, requesting, and sharing
information about the Covid-19 vaccine in a survey I received funding to
conduct of 900 Adults in the United States.

@@

\subsubsection{Article 3: Social Norms under Uncertain Times: A dynamic study of Stay-At-Home and Vaccination Rates During the Covid-19 Pandemic}
% TODO Double check name
% TODO link to article  
My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the
cases of both a) population mobility during early 2020 and b)
vaccination rates during early 2021, both as responses to public health
recommendations to mitigate the Covid-19 pandemic. Using theories of
associative diffusion and the integrated theoretical framework of norms,
I test how these health behaviors are outcomes of assortative signaling
from networks and other sources, such as social isolation and connection
between communities, religious and political conservativism, and
attention to media sources, and online norms as measured by
information-search behaviors. To investigate this, I conduct statistical
modelling on a dataset I compiled from multiple sources, including
Google, Facebook, and the Center for Disease Control and Prevention.

>>>>

A critical turning point in the study of the flow information was put forth by renowned Sociologist Paul Lazarsfeld \citeyearpar{lazarsfeldPeopleChoice1944}. Lazarsfeld and others theorized that the media did not directly influence the masses, but that a "Two-Step Flow" was occurring where "opinion leaders" interpreted messages from mass media and then, in turn, promoted their opinions to the masses \citep{katzPersonalInfluencePart1955}. The opinion leaders interpret, explain, and diffuse contents to "opinion followers". From a networks
perspective, "people rarely act on mass-media information unless it is also transmitted through personal ties" \citep[p. 1374]{granovetterStrengthWeakTies1973}. Lazarsfeld  \citeyearpar{lazarsfeldPeopleChoice1944}'s first study focused on politics but follow up studies from \citep{katzPersonalInfluencePart1955} showed that this model matched information flows beyond the political realm. In contrast to the one-step flow, where media directly influences the masses, Lazarsfeld's theory remains in prominence today. While many scholars assumed that modern micro-targeted marketing strategies would provide evidence for a one-step flow perspective \citep{bennettOneStepFlowCommunication2006}, many scholars find that "opinion leaders" (often influencers and celebrities) still create a two-step flow of information transmission \citep{choi15, hilbertOneStepTwo2017}.

@@

\subsection{Uses and Gratification Theory}

@@

\subsection{Social Norms}

@@

\section{Broader Impact}

@@

This dissertation also contributes to computational social science \&
critical big data studies. I base much of my dissertation around Google
Search Trends, which are relatively underutilized in the social sciences
compared to health sciences and business. Not only does this project
introduce Google Search Trends in new ways, but I perform explicit tests
of how this data should be used in the social sciences, providing the
groundwork for this exciting source of big data for future computational
social scientists. Because my research takes a critical big data studies
approach, I take into account the potential pitfalls with any source of
big data (McFarland and McFarland 2015) and attempt to explicitly answer
questions posed by this field, such as "what are the behavioral
processes that lead to macro-level outcomes" proposed by \citet{breigerScaling2015}.

@@

\section{Three Empirical Articles}

@@

\subsection{Article 1: Construct Validity and Correspondence of Google Trends}
% TODO Double check name
% TODO link to article  
I will first investigate how Google Search Trends can be operationalized
as cultural attitudes, disease prevalence and voting behaviors through
comparing trends data to observed valid indicators of these cases.
Specifically, I will test cultural indicators from the NORC General
Social Survey, United States county-level suicide rates from National
Center for Health Statistics, US case rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the
American National Election Survey against selected Google Trends terms
and analyze the data using correlation methods. The aim of this article
is to assess the construct validity and will contribute to critical big
data studies and sociological methodology.
% TODO ADD IN ARTICLE 1 ABSTRACT

\subsection{Article 2: Information-Seeking Behaviors for Information on Covid-19 Vaccinations}
% TODO Double check name
% TODO link to article  
The second article in my dissertation investigates the question "what
are the behavioral processes that lead to macro-level outcomes" (in this
case, Google Trends results) proposed by \citet{breigerScaling2015}. Knowing that
Google trends data only encompasses a small portion of the
information-seeking done by modern humans, it is important to
investigate what leads individuals to search for a question online using
a search engine (contributing to macro-level trends data) versus more
traditional means of information gathering, such as network activation
(asking someone in their social network for informational support). To
investigate this, I use the case of receiving, requesting, and sharing
information about the Covid-19 vaccine in a survey I received funding to
conduct of 900 Adults in the United States.

@@

\subsection{Article 3: Social Norms under Uncertain Times: A dynamic study of Stay-At-Home and Vaccination Rates During the Covid-19 Pandemic}
% TODO Double check name
% TODO link to article  
My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the
cases of both a) population mobility during early 2020 and b)
vaccination rates during early 2021, both as responses to public health
recommendations to mitigate the Covid-19 pandemic. Using theories of
associative diffusion and the integrated theoretical framework of norms,
I test how these health behaviors are outcomes of assortative signaling
from networks and other sources, such as social isolation and connection
between communities, religious and political conservativism, and
attention to media sources, and online norms as measured by
information-search behaviors. To investigate this, I conduct statistical
modelling on a dataset I compiled from multiple sources, including
Google, Facebook, and the Center for Disease Control and Prevention.

### 2022-03-02 paper1.tex
Updates

\section{Introduction}
New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
\citep{kingEnsuringDataRichFuture2011, lazerComputationalSocialScience2009,
salganikBitBitSocial2017}. Before overenthusiastically embracing these
sources into our workflows, social scientists must clearly establish
parameters under which these data sources could be operationalized
\citep{bailCulturalEnvironmentMeasuring2014, lazerParableGoogleFlu2014}. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" \citep[p. 1203]{lazerParableGoogleFlu2014}. Building on this prior research outlining these various
issues with big data \citep{boydCriticalQuestionsBig2012,
lazerIssuesConstructValidity2015}, this paper tests the construct
validity of Google Search Trends as an indicator of three different
cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators
from the NORC General Social Survey, United States county-level suicide
rates from National Center for Health Statistics, US cases rates of
Covid-19 from The New York Times, historical US Presidential Election
results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends
longitudinal data using Pearson's r for pairwise correlations, testing
for the strength of relationships between the Google Trend indicator and
the respective comparison indicator. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

I will use three categorizations of ways I propose Google Trends could
be operationalized for social scientific usage. First, I'll test Google
Trends as an operationalization of cultural attitudes with the General
Social Survey. After \citet{bailCulturalEnvironmentMeasuring2014}'s call for
cultural sociologists to utilize the ever-expanding world of big data,
Google trends as a data source began appearing in sociological and
social science research. From research on mass shootings and firearms
\citep{brownsteinInternetSearchPatterns2020,
semenzaInformationseekingWakeTragedy2020}, protest and anti-Muslim
sentiment \citep{bailUsingInternetSearch2018,
barrieSearchingRacismGeorge2020, grossThereFergusonEffect2017}, to
analyzing country-level changes in social perception \citep{reyes_etal18},
Google search trends are a new and innovative indicator of cultural
interest. Extending into social networks and culture,
\citet{bailPrestigeProximityPrejudice2019} even used Google trends to measure
how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of
disease prevalence and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic \citep{jimenez_etal20,
jimenezCOVID19SymptomGoogle2020,
limEstimatingInformationSeekingBehaviour2020,
mavraganiCOVID19PredictabilityUnited2020,
nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021} , other
research has investigated Google Trends as an indicator of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021,
carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, vaccination uptake
\citep{dalumhansenEnsembleLearnedVaccination2016}, obesity
\citep{sarigulNowcastingObesityUsing2014}, and even insomnia
\citep{zittingGoogleTrendsReveal2020}, to cover a few examples. For a partial
review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to
\citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment
\citep{dechoudhurySeekingSharingHealth2014}.

@@

\section{Methods}
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I gathered geo-located social science data across multiple
sources to address the three areas of inquiry for this paper.  Table
\@ref(tab:data-sources-table) outlines which data sources are used for this
project and which trends are matched to each source.

TABLE HERE

@@

One measure of attitudinal indicators is the Vaccine Hesitancy for COVID-19
\citep{vaches_data}. The CDC uses the U.S. Census Bureau’s Household Pulse Survey
(HPS) and the 2019 American Community Survey (ACS) 1-year Public Use Microdata
Sample (PUMS) to measure U.S. residents’ intentions to receive the COVID-19
vaccine if available during May 26, 2021 – June 7, 2021. This dataset consists
of x observations, one for each U.S. County. The variable
measures the precent of adults in the county who describe themselves as
“unsure”, “probably not”, or “definitely not” going to get a COVID-19 vaccine
once one is available to them. The variable ranges from x to y.

Another attitudinal indicator I use is the Mask-Wearing Survey Data conducted by
Dynata for the New York Times from July 2 through 14, 2020 \citep{mask_data}. 250,000
survey respondents were asked, "How often do you wear a mask in public when you
expect to be within six feet of another person?". The NYT weighted each response
to create a county level measure of what percent of the county never, rarely,
sometimes, frequently, and always wore a mask when in public. This dataset
consists of x observations, one for each U.S. County. This measure
represents the percent of adults in the county who never or rarely wear a mask
(range = x to x)

@@

I also use county-level suicide rates from the US Centers for Disease Control
and Prevention \citeyearpar{suic_data}. Data is grouped by year from 2010-2020. Raw death
rates are scaled by population size for each year and can be interpreted as the
death rate by suicide for every 1000 people. There are x total
cases, resulting from x observations of x counties. Missing data were interpolated using
`zoo::na.approx()` \citep{zoo}. Measures range from x to x.

@@

In addition to specific outcomes of interest, I also gathered various county
controls to investigate possible variable confounding. The first Seven of these
variables come from the 2010-2019 5-year American Community Surveys
\citep{acs2019, acs2018, acs2017, acs2016, acs2015, acs2014, acs2013, acs2012, acs2011, acs2010}. 
These data include total population, population density,
unemployment rate for those 16 and over, median county income, average commute
time, percent of households living under the poverty line, and percent above 65
years old. In addition, some models employ the U.S. Current Population Survey &
American Community Survey Geographic Estimates of Internet Use, 1997-2018
\citep{internet_use} to estimate households with broadband internet subscriptions.
This final variable is an attempt to capture the latent propensity to use the
internet for information search.

@@

\subsection{Analysis}
Google Search Trends data and additional demographic data are merged with each
individual indicator based on County FIPS Codes and date. After creating these
different datasets, I use the Pearson correlation formula (formula
%\@ref(eq:pearsoncorr)
) for cross-sectional numeric data to calculate the
strength of the relationship between each Google Trend and the respective
data source.

\begin{equation}
 r =
  \frac{ \sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y}) }{
        \sqrt{\sum_{i=1}^{n}(x_i-\bar{x})^2}\sqrt{\sum_{i=1}^{n}(y_i-\bar{y})^2}} %(\#eq:pearsoncorr)
\end{equation}

To address longitudinal correlations, I employ Repeated Measures Correlation
using the `rmcorr` package in R \citep{bland1995, bakdash2017}. Repeated Measures
Correlation is useful for determining the within-county association for paired
measures across time and counties. Results of these correlations can be seen in 
table %\@ref(tab:corr-results).

@@

The first attitudinal indicators is Vaccine Hesitancy for COVID-19 vaccines
\citep{vaches_data}. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4657
('Coronavirus (Disease)' and vaccine hesitancy). Correlation's under |0.40| are
considered to be weak according to the common rules of thumb. When running these
correlations in multiple linear regression (see the results in table
% \@ref(tab:vacc_hes_analysis)
), I see an $r^2$ of 0.233 (Model 1) and 0.411
(Model 2), indicating that the Google Trends are able to explain about 23% of
the variation in Vaccine Hesitancy alone. Demographic characteristics like the
percentage of households with broadband internet and the population density are
able to explain about 35% of the variation (model 2), outperforming the first
models. The Trend coefficients themselves, however, are significant and remain
significant when controlling for demographics. This reinforces the finding from
the Pearson Correlation that there is a significant but weak relationship
between the Google Trends and Vaccine Hesitancy.

The second attitudinal measure I test is how Google Trends relates to rare mask
usage. As with vaccine hesitancy, the Pearson correlations are negligible.
Correlation's under |0.20| are considered to be negigible according to the
common rules of thumb. I introduce these trends in multiple linear regression in
table 
% \@ref(tab:mask_analysis)
. Model 1 demonstrates that these five Google
Search Trends can explain about 7\% of the variance in mask usage across U.S.
counties, reinforcing the conclusion that the relationship is quite weak. The
coefficients themselves are somewhat significant in Model 1. However, after
including the demographic variables in Models 2 and 3, we see that the
relationship between the Google Search Trends and mask usage is strengthened in
magnitude and in significance, likely indicating a suppression effect due to
underlying relationships between the trends and demographic variables. While the
trends are significant and match the demographic variables in magnitude, the
low $r^2$ of 0.203 for the trends still provides evidence that Google Search
Trends data cannot replace survey analysis when trying to measure rare mask
usage.

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I test is
Covid-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of Covid-19. To further test the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table 
% \@ref(tab:covid_analysis)
indicates that the trends themselves are able
to explain about 20\% of the variation in Covid-19 rates ($r^2$ = 0.201). Few
of the demographic variables have an effect on Covid-19 rates in models 2 or 3,
though it seems there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to Covid-19 rates, but that this relationship
is weak and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

Another indicator of health outcomes I test to validate Google Trends as
possible health indicators are county suicide rates. As with previous cases,
there is only negligible correlation between actual suicide rates and Google
Search Trends, not exceeding 0.0845. In the more accurate repeated measures
correlation coefficient we see a maximum of 0.142 correlation, remaining
negligible. I introduce these trends into a random intercept model to control
for inter-group variation over time in table 
% \@ref(tab:suicide_analysis)
. While
table 
% \@ref(tab:suicide_analysis)
model 1 reveals a significant relationship
between suicide hotline trends and suicide rates, the effect is small and the
entire model only has an $r^2$ of 0.058. Adding in demographic features in model
2 improves the $r^2$ quite a lot. In model 3, the inclusion of Google Trends
actually decreases the overall amount of suicide rate variance explained
compared to model 2. These analyses lead me to conclude that Google Search
Trends should not be used as indicators of suicide rates or intentions.

@@

In table 
% \@ref(tab:pres_2016_analysis)
, I outline the results for the model
predicting the county percentage of votes for Hillary Clinton. Model 1
demonstrates how well Google Trends are able to predict the votes; while the
trend for 'Hillary Clinton' is significantly associated with votes, the model's
$r^2$ ($r^2$ = 0.027) shows that it does little to help explain the model
variance. On the other hand, adding in demographic features improves the model
fit quite well, bringing the $r^2$ up to 0.313 in model 2 and $r^2$ 0.327 in
model 3. In table 
% \@ref(tab:pres_2020_analysis)
shows this same analysis for the
2020 election and predicts the percentage of votes for Joe Biden. While this
model is largely similar to those in table 
% \@ref(tab:pres_2016_analysis)
, this
model 1 demonstrates how unrelated Google Trends can be from actual outcomes. In
conclusion, I find that Google Trends are an unreliable and invalid indicator of
political support.

>>>>

\section{Introduction}
New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
\citep{kingEnsuringDataRichFuture2011,lazerComputationalSocialScience2009,salganikBitBitSocial2017}. Before overenthusiastically embracing these
sources into our workflows, social scientists must clearly establish
parameters under which these data sources could be operationalized
\citep{bailCulturalEnvironmentMeasuring2014, lazerParableGoogleFlu2014}. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" \citep[p. 1203]{lazerParableGoogleFlu2014}. Building on this prior research outlining these various
issues with big data \citep{boydCriticalQuestionsBig2012,lazerIssuesConstructValidity2015}, this paper tests the construct
validity of Google Search Trends as an indicator of three different
cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators
from the NORC General Social Survey, United States county-level suicide
rates from National Center for Health Statistics, US cases rates of
Covid-19 from The New York Times, historical US Presidential Election
results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends
longitudinal data using Pearson's r for pairwise correlations, testing
for the strength of relationships between the Google Trend indicator and
the respective comparison indicator. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

I will use three categorizations of ways I propose Google Trends could
be operationalized for social scientific usage. First, I'll test Google
Trends as an operationalization of cultural attitudes with the General
Social Survey. After \citet{bailCulturalEnvironmentMeasuring2014}'s call for
cultural sociologists to utilize the ever-expanding world of big data,
Google trends as a data source began appearing in sociological and
social science research. From research on mass shootings and firearms
\citep{brownsteinInternetSearchPatterns2020,semenzaInformationseekingWakeTragedy2020}, protest and anti-Muslim sentiment \citep{bailUsingInternetSearch2018,barrieSearchingRacismGeorge2020,grossThereFergusonEffect2017}, to analyzing country-level changes in social perception \citep{reyes_etal18},
Google search trends are a new and innovative indicator of cultural
interest. Extending into social networks and culture, \citet{bailPrestigeProximityPrejudice2019} even used Google trends to measure
how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of
disease prevalence and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic \citep{jimenez_etal20,jimenezCOVID19SymptomGoogle2020,limEstimatingInformationSeekingBehaviour2020,mavraganiCOVID19PredictabilityUnited2020,nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021} , other
research has investigated Google Trends as an indicator of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021,carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, vaccination uptake
\citep{dalumhansenEnsembleLearnedVaccination2016}, obesity
\citep{sarigulNowcastingObesityUsing2014}, and even insomnia
\citep{zittingGoogleTrendsReveal2020}, to cover a few examples. For a partial
review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to
\citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment
\citep{dechoudhurySeekingSharingHealth2014}.

@@

\section{Methods}
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I gathered geo-located social science data across multiple
sources to address the three areas of inquiry for this paper.  Table
\ref{tab:data-sources-table} outlines which data sources are used for this
project and which trends are matched to each source.

\input{tables/paper1/data-sources-table}

@@

One measure of attitudinal indicators is the Vaccine Hesitancy for COVID-19
\citep{vaches_data}. The CDC uses the U.S. Census Bureau’s Household Pulse Survey
(HPS) and the 2019 American Community Survey (ACS) 1-year Public Use Microdata
Sample (PUMS) to measure U.S. residents’ intentions to receive the COVID-19
vaccine if available during May 26, 2021 – June 7, 2021. This dataset consists
of x observations, one for each U.S. County. The variable
measures the percent of adults in the county who describe themselves as
“unsure”, “probably not”, or “definitely not” going to get a COVID-19 vaccine
once one is available to them. The variable ranges from x to y.

Another attitudinal indicator I use is the Mask-Wearing Survey Data conducted by
Dynata for the New York Times from July 2 through 14, 2020 \citep{mask_data}. 250,000
survey respondents were asked, "How often do you wear a mask in public when you
expect to be within six feet of another person?". The NYT weighted each response
to create a county level measure of what percent of the county never, rarely,
sometimes, frequently, and always wore a mask when in public. This dataset
consists of x observations, one for each U.S. County. This measure
represents the percent of adults in the county who never or rarely wear a mask
(range = x to x).

@@

I also use county-level suicide rates from the US Centers for Disease Control
and Prevention \citeyearpar{suic_data}. Data is grouped by year from 2010-2020. Raw death
rates are scaled by population size for each year and can be interpreted as the
death rate by suicide for every 1000 people. There are x total
cases, resulting from x observations of x counties. Missing data were interpolated using
zoo::na.approx() \citep{zoo}. Measures range from x to x.

@@

In addition to specific outcomes of interest, I also gathered various county
controls to investigate possible variable confounding. The first Seven of these
variables come from the 2010-2019 5-year American Community Surveys
\citep{acs2019, acs2018, acs2017, acs2016, acs2015, acs2014, acs2013, acs2012, acs2011, acs2010}. 
These data include total population, population density,
unemployment rate for those 16 and over, median county income, average commute
time, percent of households living under the poverty line, and percent above 65
years old. In addition, some models employ the U.S. Current Population Survey \&
American Community Survey Geographic Estimates of Internet Use, 1997-2018
\citep{internet_use} to estimate households with broadband internet subscriptions.
This final variable is an attempt to capture the latent propensity to use the
internet for information search.

@@

\subsection{Analysis}
Google Search Trends data and additional demographic data are merged with each
individual indicator based on County FIPS Codes and date. After creating these
different datasets, I use the Pearson correlation formula (formula \eqref{eq:pearsoncorr}
) for cross-sectional numeric data to calculate the
strength of the relationship between each Google Trend and the respective
data source.

\begin{equation}
 r =
  \frac{ \sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y}) }{
        \sqrt{\sum_{i=1}^{n}(x_i-\bar{x})^2}\sqrt{\sum_{i=1}^{n}(y_i-\bar{y})^2}} \label{eq:pearsoncorr}
\end{equation}

To address longitudinal correlations, I employ Repeated Measures Correlation
using the `rmcorr` package in R \citep{bland1995, bakdash2017}. Repeated Measures
Correlation is useful for determining the within-county association for paired
measures across time and counties. Results of these correlations can be seen in 
table \ref{tab:corr-results}.

@@

\input{tables/paper1/corr-results}

The first attitudinal indicators is Vaccine Hesitancy for COVID-19 vaccines
\citep{vaches_data}. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4657
('Coronavirus (Disease)' and vaccine hesitancy). Correlation's under |0.40| are
considered to be weak according to the common rules of thumb. When running these
correlations in multiple linear regression (see the results in table
\ref{tab:vacc_hes_analysis}), I see an $r^2$ of 0.233 (Model 1) and 0.411
(Model 2), indicating that the Google Trends are able to explain about 23% of
the variation in Vaccine Hesitancy alone. Demographic characteristics like the
percentage of households with broadband internet and the population density are
able to explain about 35\% of the variation (model 2), outperforming the first
models. The Trend coefficients themselves, however, are significant and remain
significant when controlling for demographics. This reinforces the finding from
the Pearson Correlation that there is a significant but weak relationship
between the Google Trends and Vaccine Hesitancy.

\input{tables/paper1/vacc_hes_analysis}

The second attitudinal measure I test is how Google Trends relates to rare mask
usage. As with vaccine hesitancy, the Pearson correlations are negligible.
Correlation's under |0.20| are considered to be negigible according to the
common rules of thumb. I introduce these trends in multiple linear regression in
table \ref{tab:mask_analysis}. Model 1 demonstrates that these five Google
Search Trends can explain about 7\% of the variance in mask usage across U.S.
counties, reinforcing the conclusion that the relationship is quite weak. The
coefficients themselves are somewhat significant in Model 1. However, after
including the demographic variables in Models 2 and 3, we see that the
relationship between the Google Search Trends and mask usage is strengthened in
magnitude and in significance, likely indicating a suppression effect due to
underlying relationships between the trends and demographic variables. While the
trends are significant and match the demographic variables in magnitude, the
low $r^2$ of 0.203 for the trends still provides evidence that Google Search
Trends data cannot replace survey analysis when trying to measure rare mask
usage.

\input{tables/paper1/mask_analysis}

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I test is
Covid-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of Covid-19. To further test the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table \ref{tab:covid_analysis} indicates that the trends themselves are able
to explain about 20\% of the variation in Covid-19 rates ($r^2$ = 0.201). Few
of the demographic variables have an effect on Covid-19 rates in models 2 or 3,
though it seems there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to Covid-19 rates, but that this relationship
is weak and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

\input{tables/paper1/covid_analysis}

Another indicator of health outcomes I test to validate Google Trends as
possible health indicators are county suicide rates. As with previous cases,
there is only negligible correlation between actual suicide rates and Google
Search Trends, not exceeding 0.0845. In the more accurate repeated measures
correlation coefficient we see a maximum of 0.142 correlation, remaining
negligible. I introduce these trends into a random intercept model to control
for inter-group variation over time in table \ref{tab:suicide_analysis}. While
table \ref{tab:suicide_analysis} model 1 reveals a significant relationship
between suicide hotline trends and suicide rates, the effect is small and the
entire model only has an $r^2$ of 0.058. Adding in demographic features in model
2 improves the $r^2$ quite a lot. In model 3, the inclusion of Google Trends
actually decreases the overall amount of suicide rate variance explained
compared to model 2. These analyses lead me to conclude that Google Search
Trends should not be used as indicators of suicide rates or intentions.

\input{tables/paper1/suicide_analysis}

@@

In table \ref{tab:pres_2016_analysis}, I outline the results for the model
predicting the county percentage of votes for Hillary Clinton. Model 1
demonstrates how well Google Trends are able to predict the votes; while the
trend for 'Hillary Clinton' is significantly associated with votes, the model's
$r^2$ ($r^2$ = 0.027) shows that it does little to help explain the model
variance. On the other hand, adding in demographic features improves the model
fit quite well, bringing the $r^2$ up to 0.313 in model 2 and $r^2$ 0.327 in
model 3. In table \ref{tab:pres_2020_analysis} shows this same analysis for the
2020 election and predicts the percentage of votes for Joe Biden. While this
model is largely similar to those in table \ref{tab:pres_2016_analysis}, this
model 1 demonstrates how unrelated Google Trends can be from actual outcomes. In
conclusion, I find that Google Trends are an unreliable and invalid indicator of
political support.

\input{tables/paper1/pres_2016_analysis}
\input{tables/paper1/pres_2020_analysis}
